### 2019-09-17 pres/moltres.tex
Last paper changes and first presentation commit

\begin{frame}
	\frametitle{Moltres}
		\textbf{What is Moltres?}
		\begin{itemize}
			\item Moltres is an application code built on the \gls{MOOSE}
			framework, for the simulation of \glspl{MSR}
			\item \gls{MOOSE} is an open source finite element framework that
			relies on Libmesh and PETSc for advanced meshing and PDE solving
			capabilities
			\item Moltres can run transient, implicitly coupled
			neutronics/thermal-hydraulics simulations
			\begin{itemize}
				\item Multi-group neutron diffusion (arbitrary groups)
				\item Delayed neutron precursor decay (with advection)
				\item Incompressible Navier-Stokes for temperature
				advection-diffusion
				\item 1D, 2D, and 3D modeling are supported
			\end{itemize}
		\end{itemize}
\end{frame}