### 2019-09-19 pres/conclusion.tex
presentation slides first draft

\begin{frame}
	\frametitle{Conclusion}
		The neutron group fluxes and reactivity coefficients from Moltres
		(without \gls{DNP} drift) are in
		good agreement with Serpent results, in the context of the \gls{MSFR}.

		\vspace{.3cm}
		The steady state neutron flux, temperature, and \gls{DNP} distributions
		observed are expected for uniform flow.

		\vspace{.3cm}
		Discrepancies in peak neutron
		flux and temperature distributions to other work are primarily due to
		the minor differences in core geometry and using uniform flow instead of
		appling a turbulence model.

		\vspace{.3cm}
		Higher core temperatures observed for start-up fuel composition than
		early-life and equilibrium compositions.

		\vspace{.3cm}
		Discrepancy in average core temperatures during \gls{ULOHS} is mainly due
		to the lack of a decay heat model in Moltres.
\end{frame}