### 2019-09-19 pres/objectives.tex

\begin{frame}
	\frametitle{Objectives}
		\textbf{Objectives}
		\begin{itemize}
			\item Verify the neutronics capabilities in Moltres against Serpent
			\cite{leppanen_serpent_2015}
			\item Demonstrate coupled neutronics/thermal-hydraulics simulations
			of the \gls{MSFR} concept,
			\begin{itemize}
				\item with start-up, early-life, and equilibrium fuel
				compositions
				\item for steady state and transient cases: \gls{ULOHS} accident
			\end{itemize}
		\end{itemize}
\end{frame}