### 2019-10-12 pres/acks.tex
Final version

\begin{frame}
	\frametitle{Acknowledgments}
    	Sun Myung Park is supported by the Singapore Nuclear Research and Safety
    	Initiative (SNRSI) Postgraduate Scholarship.

    	\vspace{.3cm}
    	Andrei Rykhlevskii and Dr Kathryn Huff are both supported by funding from
    	the DOE ARPA-E MEITNER Program (award DE-AR0000983).

    	\vspace{.3cm}
    	Dr Kathryn Huff is also supported by the Blue Waters sustained-petascale
    	computing project supported by the National Science Foundation.
\end{frame}

\begin{frame}
	\large
	\centering
	\textbf{Thank you for your attention!}
\end{frame}