### 2019-10-12 pres/conclusion.tex

\begin{frame}
	\frametitle{Conclusion}
		The neutron group fluxes and reactivity coefficients from Moltres
		(without \gls{DNP} drift) are in
		good agreement with Serpent results.

		\vspace{.3cm}
		The steady state neutron flux, temperature, and \gls{DNP} distributions
		observed are expected for uniform flow.

		\vspace{.3cm}
		Discrepancies present in peak neutron
		flux and temperature distributions relative to other work

		\vspace{.3cm}
		Higher core temperatures observed for start-up fuel composition than
		early-life and equilibrium compositions.
\end{frame}