### 2019-10-12 pres/objectives.tex

\begin{frame}
	\frametitle{Objectives}
		\textbf{Objectives}
		\begin{itemize}
			\item Verify the neutronics results in Moltres against Serpent
			\cite{leppanen_serpent_2015} with the \gls{MSFR} concept with six
			neutron energy groups
			\begin{itemize}
				\item Previous validation study has been performed with an
				MSRE-like model with two (fast/thermal) neutron energy groups
				\cite{lindsay_introduction_2018}
			\end{itemize}
			\item Demonstrate coupled neutronics/thermal-hydraulics simulations
			of the \gls{MSFR} concept using Moltres,
			\begin{itemize}
				\item with start-up, early-life, and equilibrium fuel
				compositions
				\item for steady state and transient cases: \gls{ULOHS} accident
			\end{itemize}
		\end{itemize}
\end{frame}